Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex
Abstract tweak.
more copyediting

Verified systems have been growing in popularity recently with the rise of proof
assistants like Coq and NuPRL~\cite{coqart,nuprl}.  
These systems allow programmers to construct proofs of properties of their programs which are checked and compiled by the system.  As we will
discuss in Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to formalize,
prove correct, and implement systems with verified properties.  This increases Programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).

@@

Given the growing popularity of Proof Assistants information-flow languages, we were surprised to discover 
that there has not been work done to formalize ad verify a language with a type systems which enforces information flow security policies.
These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties it enforces.
Our goal is to make progress towards verifying the correctness of a
compiler for a language with information flow types that programmers will want
to use.

The core theorem which distinguishes an information-flow language from a 
``normal'' programming language is the poof of noninterference, 
which asserts that across any two runs, every program in the language will never
produce different outputs for low security items if only the high security input
values are varied.  It is just this theorem we wish to prove in this paper; we
hope to learn through this effort of the unique challenges associated with
mechanizing proofs of information-flow languages.

Our language, \langName,  and proof are a simplified version of what is
presented in Simonet and Pottier's work on information flow for an ML-like
language, \origLang~\cite{InfoFlowML}.  The proof proceeds by designing a
core language, extending it with a term that represents a ``pair'' of 
values varied across two executions, and introducing these ``pairs'' of 
different values in context where the value would have a high-security label.
The proof then becomes a matter of showing that if
the program is typed with a low security label, it does not terminate
with a result that is an instance of one of these ``pairs.''

@@

\begin{itemize}
  \item References and associated memory operations.
  \item Value Tuples.
  \item Exceptions.
  \item Sequencing operations.
\end{itemize}

While the language becomes much more burdensome for the programmer, we
feel that these features do not fundamentally change the problem of proving
noninterference.  Mechanizing a proof of noninterference for a
language closer to \origLang would require simply adding mechanized proofs
for the additional cases where appropriate.

>>>>

Verified systems have been growing in popularity recently with the rise of proof
assistants like Coq and NuPRL~\cite{coqart,nuprl}.  
These systems allow programmers to construct proofs of properties of their programs which are checked and compiled by the system.  As we will
discuss in Section~\ref{sec:related}, these \emph{Proof Assistants} have been used to formalize,
prove correct, and implement systems with verified properties.  This increases programmers' confidence in their languages and programs; to prove correctness of a function is a much stronger guarantee than simply testing functionality (or asserting correctness in the absence of tests).

@@

Given the growing popularity of Proof Assistants information-flow languages, we were surprised to discover 
that there has not been work done to formalize ad verify a language with a type system which enforces information flow security policies.
These languages seem a natural choice for formal verification, as it is desirable for programs written in these languages to be accompanied with a proof of the security properties they enforce.
Our goal is to make progress towards verifying the correctness of a
compiler for a language with information flow types that programmers will want
to use.

The core theorem which distinguishes an information-flow language from a 
``normal'' programming language is the poof of noninterference, 
which asserts that across any two runs, every program in the language will never
produce different outputs for low security items if only the high security input
values are varied.  It is this theorem we wish to prove in this paper; we
hope to learn through this effort of the unique challenges associated with
mechanizing proofs of information-flow languages.

Our language and proof, \langName, are a simplified version of what is
presented in Simonet and Pottier's work on information flow for an ML-like
language, \origLang~\cite{InfoFlowML}.  The proof proceeds by designing a
core language, extending it with a term that represents a ``pair'' of 
values varied across two executions, and introducing these ``pairs'' of 
values in contexts where the value would have a high-security label.
The proof then becomes a matter of showing that if
the program is typed with a low security label, it does not terminate
with a result that is an instance of one of these ``pairs.''

@@

\begin{itemize}
  \item References and associated memory operations.
  \item Value 2-Tuples.
  \item Exceptions.
  \item Sequencing operations.
\end{itemize}

While programming in this language becomes much more burdensome, we
feel that the removal of these features does not fundamentally 
change the problem of proving
noninterference.  Mechanizing a proof of noninterference for a
language closer to \origLang would require simply adding mechanized proofs
for the additional cases where appropriate.
